---
title: "140.615.HW.07.Jin.Vincent"
author: "Vincent Jin"
date: "2023-04-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 07
## Vincent Jin

# 1. 
Two drugs, zidovudine (Z) and didanosine (D), were tested for their effectiveness in preventing progression of HIV disease in children. In a double-blind clinical trial, 276 children with HIV were given Z, 281 were given D, and 274 were given both Z and D. The following table shows the survival data for the three groups.
$$ \begin{matrix}
Treatment & Died & Survived & Total \\
Z & 17 & 259 & 276 \\
D & 7 & 274 & 281 \\
Z+D & 10 & 264 & 274 
\end{matrix}$$
We want to test whether survival and treatment are independent.

## (a)
Calculate the chi-square test statistic and a corresponding p-value. Confirm your findings using the R function chisq.test().

``` {r}
x <- rbind(c(17, 259), c(7, 274), c(10, 264))
rs <- apply(x, 1, sum) 
cs <- apply(x, 2, sum) 
n  <- sum(x)
e <- outer(rs, cs, "*") / n
xsq <- sum((x - e)^2/e)
df <- (2 - 1) * (3 - 1)
print(paste("The chi-square statistic is", xsq))
print(paste("The p-value associated with a chi-square statistic of", xsq, "under a degree of freedom of 4 is:"))
pchisq(xsq, df, lower.tail=FALSE)
print("The result from chisq.test() function is:")
chisq.test(x)
```

## (b)
Calculate the likelihood ratio test statistic and a corresponding p-value.

``` {r}
lrt <- 2 * sum(x * log(x / e)) 
print(paste("The likelihood ratio test statistic is", lrt))
print(paste("The p-value associated with a likelihood ratio test statistic of", lrt, "is"))
pchisq(lrt, df, lower.tail = FALSE)
```

## (c)
Perform Fisher’s exact test

``` {r}
fisher.test(x)
```

## (d)
What do you conclude?

***Answer***

The chi-sqaure, likelihood ratio test and Fisher's exact test all suggested p-values that are greater than 0.05 （0.083, 0.087, 0.084, repectively). Therefore, we failed to reject the null hypothesis of susvival and treatments are independent.

# 2.
A  researcher studied patients with Clostridium difficile infections, which cause persistent diarrhea. Some patients were given the antibiotic vancomycin, and some patients were given a fecal transplant. Among the people who received a fecal transplant 13 out of 16 were cured, while only 4 out of 13 patients who received vancomycin were cured. Use Fisher’s exact test to test the null hypothesis that there is no difference in treatment versus the alternative that the fecal transplant is better than vancomycin. Do so by calculating the p-value based on the respective hypergeometric probabilities, and confirm your result
using the built-in R function.

***Answer***

Based on the information provided, we can construct a table of treatment vs. cure like below:
$$\begin{matrix}
\ & Cured & Uncured \\
Fecal\ Transplant & 13 & 3 \\
Vancomycin & 4 & 9
\end{matrix}$$





