---
title: "140.615.HW.07.Jin.Vincent"
author: "Vincent Jin"
date: "2023-04-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 07
## Vincent Jin

# 1. 
Two drugs, zidovudine (Z) and didanosine (D), were tested for their effectiveness in preventing progression of HIV disease in children. In a double-blind clinical trial, 276 children with HIV were given Z, 281 were given D, and 274 were given both Z and D. The following table shows the survival data for the three groups.
$$ \begin{matrix}
Treatment & Died & Survived & Total \\
Z & 17 & 259 & 276 \\
D & 7 & 274 & 281 \\
Z+D & 10 & 264 & 274 
\end{matrix}$$
We want to test whether survival and treatment are independent.

## (a)
Calculate the chi-square test statistic and a corresponding p-value. Confirm your findings using the R function chisq.test().

``` {r}
x <- rbind(c(17, 259), c(7, 274), c(10, 264))
rs <- apply(x, 1, sum) 
cs <- apply(x, 2, sum) 
n  <- sum(x)
e <- outer(rs, cs, "*") / n
xsq <- sum((x - e)^2/e)
df <- (2 - 1) * (3 - 1)
print(paste("The chi-square statistic is", xsq))
print(paste("The p-value associated with a chi-square statistic of", xsq, "under a degree of freedom of 4 is:"))
pchisq(xsq, df, lower.tail=FALSE)
print("The result from chisq.test() function is:")
chisq.test(x)
```

## (b)
Calculate the likelihood ratio test statistic and a corresponding p-value.

``` {r}
lrt <- 2 * sum(x * log(x / e)) 
print(paste("The likelihood ratio test statistic is", lrt))
print(paste("The p-value associated with a likelihood ratio test statistic of", lrt, "is"))
pchisq(lrt, df, lower.tail = FALSE)
```

## (c)
Perform Fisher’s exact test

``` {r}
fisher.test(x)
```




